Expert-In-The-Loop Causal Discovery: Iterative Model Refinement Using Expert Knowledge

### 2025-02-10 paper/main.tex
Updates

\begin{algorithm}[h]
\DontPrintSemicolon
\SetAlgoLined
\SetKwFunction{RankedExpand}{RankedExpand}
\SetKw{KwGoTo}{go to}
\SetKwProg{Fn}{Function}{:}{}
\Fn{{\sc RankedExpand}($V,E,\mathcal{D}_G,\mathcal{A}_G$,$\phi$,$B=\emptyset$)}{
    let $\phi_{\max} \gets 0 $\;
    \ForEach{$X, Y$ where $X \to Y \notin E$ and $Y \to X \notin E$ and 
	$\{X,Y\} \notin B$}{
        let $\mathbf{Z}$ be a set that $d$-separates $X$ and $Y$ in $(V,E)$\;
        \If{$\mathcal{D}_G(X, Y, \mathbf{Z}) = 0$}{
		\If{$\lvert \phi(X, Y, \mathbf{Z}) \rvert > \phi_{\max}$}{
			$ \phi_{\max} \gets \lvert \phi(X, Y, \mathbf{Z}) \rvert $ \;
			$ L_{\max} \gets \mathcal{A}_G(X, Y)$ \;
		}
        }
    }
    add $L_{\max}$ to $E$\;
    \Return{$(V,E)$}\;
}
\caption{Adding an edge between variables with the highest correlation}
\label{algo:ranked_expand}
\end{algorithm}

@@

Therefore, an alternative to selecting an edge at random is to rank potential
edges based on the residual association between the variables given the DAG.
For quantifying the residual association between variables, we can use the
effect size measures of the CI test. So, for a d-separation query $ D_G(X, Y,
\mathbf{Z}) $, we can quantify the residual association between $ X $ and $ Y $
conditioned on $ \mathbf{Z} $ as the effect size of the CI test $ X \ci Y |
\mathbf{Z} $. A greedy approach could be to choose the edge that has the
highest residual association between the variables.
Algorithm~\ref{algo:ranked_expand} outlines this approach. This approach has
the benefit that it prioritizes edges that improve the model fit the most.

>>>>

\begin{algorithm}[h]
\DontPrintSemicolon
\SetAlgoLined
\SetKwFunction{RankedExpand}{RankedExpand}
\SetKw{KwGoTo}{go to}
\SetKwProg{Fn}{Function}{:}{}
\Fn{{\sc RankedExpand}($V,E,\mathcal{D}_G,\mathcal{A}_G$,$\phi$,$B=\emptyset$)}{
    let $\phi_{\max} \gets 0 $\;
    \ForEach{$X, Y$ where $X \to Y \notin E$ and $Y \to X \notin E$ and 
	$\{X,Y\} \notin B$}{
        let $\mathbf{Z}$ be a set that $d$-separates $X$ and $Y$ in $(V,E)$\;
        \If{$\mathcal{D}_G(X, Y, \mathbf{Z}) = 0$}{
		\If{$\lvert \phi(X, Y, \mathbf{Z}) \rvert > \phi_{\max}$}{
			$ \phi_{\max} \gets \lvert \phi(X, Y, \mathbf{Z}) \rvert $ \;
			$ e_{\max} \gets \mathcal{A}_G(X, Y)$ \;
		}
        }
    }

    \uIf{ $(V,E \cup e_{\max})$ is acyclic }{
	add $ e_{\max}$ to $E$ \;
    } \Else {
	add $ \textrm{reverse}(e_{\max})$ to $E$ \;
    }
    \Return{$(V,E)$}\;
}
\caption{Adding an edge between variables with the highest correlation}
\label{algo:ranked_expand}
\end{algorithm}

@@

Therefore, an alternative to selecting an edge at random is to rank potential
edges based on the residual association between the variables given the DAG.
For quantifying the residual association between variables, we can use the
effect size measures of the CI test. So, for a d-separation query $ D_G(X, Y,
\mathbf{Z}) $, we can quantify the residual association between $ X $ and $ Y $
conditioned on $ \mathbf{Z} $, $ \phi(X, Y, \bm{Z}) $ as the effect size of the
CI test $ X \ci Y | \mathbf{Z} $. A greedy approach could be to choose the edge
that has the highest absolute residual association between the variables.
Algorithm~\ref{algo:ranked_expand} outlines this approach. This approach has
the benefit that it prioritizes edges that improve the model fit the most. We use
the \textsc{RankedExpand} in all of our experiments in the next section.
